---
title: "Water Quality Time Series Analysis"
author: "Brandon Bagley"
output:
  html_document: default
  html_notebook: default
---
```{r setup, message=F, echo = F}
suppressWarnings(suppressMessages(library(tidyverse)))
suppressWarnings(suppressMessages(library(PerformanceAnalytics)))
suppressWarnings(suppressMessages(library(ISLR)))
suppressWarnings(suppressMessages(library(imputeTS)))
set.seed(10)
# Read in raw file
raw_4870 <- read.csv('data/USGS_03434870_water', header = T, skip = 31, sep = '\t')
raw_4870 <- raw_4870[-1,] # Remove first line of nonsense
```

```{r munging_cumberland, echo = F}
wat_cumb <- raw_4870 %>% 
  select(agency_cd, site_no, datetime, temp_c_av=X130015_00010_00003, 
         SpC_ms_av=X130016_00095_00003, pH_av=X130017_00400_00003, 
         DO_mgl_av=X130018_00300_00003) %>% # Rename based on metadata
  mutate(site_no = paste0(as.character(agency_cd),'_',as.character(site_no)),
         datetime = as.Date.character(datetime, "%Y-%m-%d"),
         temp_c_av = as.numeric(as.character(temp_c_av)),
         SpC_ms_av = as.numeric(as.character(SpC_ms_av)),
         pH_av = as.numeric(as.character(pH_av)), 
         DO_mgl_av = as.numeric(as.character(DO_mgl_av)),
         random = rbinom(nrow(.), 2, .05))
```

Data for surface water was downloaded from the [USGS Water Data For the Nation](https://waterdata.usgs.gov/nwis) website for a site in Tennessee that had multiple measurements for at least a year.  The data is used here to simulate a year of readings from another source to try and identify treands and predict a randomly assigned variable.  The randomly assigned variable could represent a filter change, or maintenance which in the data is named `random`.  The random variable was assigned to remain 0 for 95% of the timespan of the data which ranged from `r min(wat_cumb$datetime)` to `r max(wat_cumb$datetime)`, and score a 1 or 2 for only 5% of the timespan.  

The variables measured at the USGS site #03434870 from the river station were:

+ Temperature (Celsius)   
`temp_c_av`
+ Specific Conductance (Ms per cm at 25C)  
`SpC_ms_av`
+ pH (Field)  
`pH_av`
+ Dissolved Oxygen (Mg/L)  
`DO_mgl_av`
+ Random (Simulating rare occurance event)  
`random`

## Data exploration
```{r plot_all, echo = F}
plot_dta <- wat_cumb %>% 
  select(datetime, temp_c_av, SpC_ms_av, pH_av, DO_mgl_av, random) %>% 
  gather("parameter", "value", -datetime) %>% 
  na.omit()

ggplot(plot_dta, aes(x = datetime, y = value, color = parameter)) +
  geom_point(size = .5) +
  scale_x_date(date_breaks = '4 months', date_labels = '%b-%Y')+
    theme_light() +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 30, hjust = 1, size = 8)) +
  facet_wrap(~parameter, scales = "free_y") +
  labs(title = "Exploratory analysis at the Cumberland River, TN",
       subtitle = "All Paramaters plotted through time",
       x = "")

```
```{r}
wat_seq <- seq(wat_cumb$datetime[1], wat_cumb$datetime[nrow(wat_cumb)], by = "day")
wat_sq <- c(2015, as.numeric(format(wat_seq[1], "%j")))

wa_im <- ts(wat_cumb[,c(4:7)], start = wat_sq,
            frequency = 365)

wa_imp <- na.kalman(wa_im, model = "StructTS")
plotNA.gapsize(wa_im, wa_imp)
```

The data is showing seasonal trends in alomst all of the measured paramaters which is to be expected.  There appears to be a strong seasonal trean in the temperature `temp_c_av`, but no seasonal trend in the random data.  A closer examination of the data shows a sine-like wave trend in the temperature that is well fitted.


```{r exploration, echo = F}
Time <- as.numeric(wat_cumb$datetime)
temp <- wat_cumb$temp_c_av

xc <- cos(2*pi*Time/366)
xs <- sin(2*pi*Time/366)
fit.lm <- lm(temp~xc+xs)
fit <- fitted(fit.lm)

pred <- predict(fit.lm, newdata = data.frame(Time=time))

plot(temp_c_av~as.numeric(datetime), data = wat_cumb,
     xlab = 'Time', ylab = "Temperature(C)", pch = 19, cex = .5)
lines(Time, pred, col="blue")
title(main = "Water Temperature Through Time at the Cumberland River, TN\nSinesoidal Predicted wave overlain")
```
```{r cumb_temp, eval = F, echo = F}
# ggplot(wat_cumb, aes(x = as.numeric(datetime), y = temp_c_av)) +
#   geom_point(color = 'grey40', size = .5)+
#   geom_line(color = 'blue', data = pred, aes(x=Time, y=as.numeric(datetime)))+
#   scale_x_date(date_breaks = '3 months', date_labels = '%b-%Y')+
#   labs(title = "Water Temperature (C) at USGS_03434870",
#        x = "", y = "Temperature C") +
#   theme_light()
```


To further explore the relationship between the data and their correlation to one antoher, a correlation plot with a histogram and significance overlays can be completed using the _Spearman_ correlation coefficent to identify linear and non-linear trends.

```{r corr, echo = F}
plot_dta = wat_cumb %>% 
  mutate(date = as.numeric(as.POSIXct(wat_cumb$datetime, format = "%Y-%m-%d"))) %>% 
  select(date, temp_c_av, SpC_ms_av, pH_av, DO_mgl_av, random)
correlation <- cor(plot_dta, method = "spearman", use = "complete.obs")
chart.Correlation(plot_dta, histogram = T, pch = 19, method = 'spearman')
```


The plots indicate a strong relationship between Dissolved oxygen (DO) and temperature.


## First Steps

1. Scale the data
2. identify correlation among scaled results
3. Determine training/test/validation datasets via sampling techinques
+ Could be based on a single "pump" vs all other pumps or percentage split, also resampling(bootstrap)
4. build model
5. Interogate results and if possible the residuals
+ Identify Accuracy, but also ROC curves (AUC)
6. Re-evaluate model
7. Model validation via CV, OOB, Resampling techinques

## Models
### Logistic Regression to determine function of random from inputs
### Support Vector Machine to identify linear or using a kernel- non-linear function to distinguish between 0 & 1 assuming that 2 is within 1
### Neural Network to determine random variable with learning based on the probability of classes (0,1,2)
### Boosted Random FOrest to train the model on outliers whcih is the rare instances of 1 or 2 random


```{r}
wat_max <- apply(wat_cumb[,3:8], 2, max)
wat_min <- apply(wat_cumb[,3:8], 2, min)
# wat_scaled <- as.data.frame(scale(wat_cumb[,3:8], center = wat_min, scale = wat_max - wat_min))
fft(wat_cumb[,3:8])

```

